%ws1.solns.tex
%COMS10014 
%taught at the University of Bristol
%2020_21 Conor Houghton conor.houghton@bristol.ac.uk

%To the extent possible under law, the author has dedicated all copyright 
%and related and neighboring rights to these notes to the public domain 
%worldwide. These notes are distributed without any warranty. 

\documentclass[11pt,a4paper]{scrartcl}
\typearea{12}
\usepackage{graphicx}
%\usepackage{pstricks}
\usepackage{listings}
\usepackage{color}
\lstset{language=C}
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{\texttt{cs-uob.github.io/COMS10014/ and github.com/coms10011/2020\_21}}
\lfoot{COMS10014 - P\&C ws3 solns - Conor}
\begin{document}

\section*{Worksheet 3 - outline solutions}

\subsection*{Questions}

\begin{enumerate}

\item Two events $A$ and $B$ have probabilities $P(A)=0.2$, $P(B)=0.3$ and $P(A\cup B)=0.4$. Find
\begin{enumerate}
\item Find $P(A\cap B)$.
\item Find $P(\bar{A}\cap \bar{B})$.
\item Find $P(A|B)$.
\end{enumerate}
\textbf{Solution}: So
\begin{equation}
P(A\cup B)=P(A)+P(B)-P(A\cap B)
\end{equation}
so 
\begin{equation}
P(A\cap B)=0.2+0.3-0.4=0.1
\end{equation}
If $P(A\cap B)=0.1$ then $P(\bar{A}\cap B)=P(B)-P(A\cap B)=0.2$. One the other hand
$P(\bar{A}\cap \bar{B})=1-P(A\cup B)=0.6$. Finally
\begin{equation}
P(A|B)=\frac{P(A\cap B)}{P(B)}=\frac{0.1}{0.3}=\frac{1}{3}
\end{equation}


\item In a library where all books have blue or yellow spines, four
  fifths of books with yellow spines are about mathematics but only a
  fifth of books with blue spines are about mathematics. There are the
  same number of yellow and blue spined books, you come upom a book
  open on a table; the book is about mathematics. What is the chance
  it has a yellow spine?\\ \\ \\
  \textbf{Solution}: In the obvious notation we need $P(M)$:
  $$P(M)=P(M|B)P(B)+P(M|Y)P(Y)=0.1+0.4=0.5$$
  and using Bayes:
  $$P(Y|M)=pP(M|Y)P(Y)/P(M)=0.4/0.5=0.8$$  


\item You want to go for a walk. However, when you wake up the day is
  cloudy and half of all raining days start off cloudy. On the other
  hand, two days in five start off cloudy and it's been rather dry
  recently with only rain only on one day in ten. What is the chance
  it will rain?\\ \\ \\
  \textbf{Solution}:
  \begin{equation}
      P(\mbox{rain}|\mbox{cloudy})=\frac{P(\mbox{cloudy}|\mbox{rain})P(\mbox{rain})}{P(\mbox{cloudy})}=\frac{0.5*0.1}{0.4}=0.125
    \end{equation}


  
\item One night in a bar in Las Vegas you meet a dodgy character who
  tells you that there are two types of slot machine in the Topicana,
  one that pays out 10\% of the time, the other 20\%. One sort of
  machine is blue, the other red. Unfortunately the dodgy character is
  too drunk to remember which is which. The next day you randomly
  select red to try, you find a red machine and put in a coin. You
  lose. Assuming the dodgy character was telling the truth, what is
  the chance the red machine is the one that pays out more. If you had
  won instead of losing, what would the chance be?\footnote{I stole
    this problem from \texttt{courses.smp.uq.edu.au/MATH3104/}}
\\ \\ \\ \textbf{Solution}:
  So let $J$ be the event of winning and $R$ be the event that the red
  machine is the one that pays out more. We want
\begin{equation}
P(R|!J)=\frac{P(!J|R)P(R)}{P(!J)}
\end{equation}
where we are writing $!J$, not $J$, for $\bar{J}$. Now $P(!J|R)=0.8$
according to the DC and $P(R)=0.5$ because we chose between red and
blue randomly. Finally
\begin{equation}
P(!J)=P(!J|R)P(R)+P(!J|B)P(B)=0.8\times 0.5 + 0.9\times 0.5=0.85
\end{equation}
and hence
\begin{equation}
P(R|\bar{J})=\frac{0.8\times 0.5}{0.85}\approx0.47
\end{equation}
so, consider the prior $P(R)=0.5$ we haven't learned much from this lost coin. If we had won you'd have
\begin{equation}
P(R|J)=\frac{P(J|R)P(R)}{P(J)}=\frac{0.2\times 0.5}{0.15}\approx 0.67
\end{equation}
so you'd learn a lot more from a win, this makes sense since it is a rarer event.
\end{enumerate}

  
\subsection*{Some more questions}

\begin{enumerate}
\item $A=\{a,b,c\}$, $B=\{c,d,e\}$ and the sample space $X=A\cup B\cup \{f,g,h,i\}$. All outcomes, that is the letters, are equally likely. What is $P(A)$? What is $P(A|B)$? Are $A$ and $B$ independent?\\ \\ \\
  \textbf{Solution}: So $P(A)=3/9=1/3$ and
  \begin{equation}
    P(A|B)=\frac{P(A\cap B)}{P(B)}=\frac{1}{3}
  \end{equation}
  so $A$ and $B$ are independent.

\item $A=\{a,b,c,j,k\}$, $B=\{c,d,e,l\}$, $C=\{a,b,c,d,e,f,g,h,i\}$ and
  $X=A\cup B\cup C\cup \{m,n\}$. Are $A$ and $B$ independent given $C$? Are they
  conditionally independent?\\ \\ \\ \textbf{Solution}: well
  $P(A)=5/14$ and $P(B)=4/14$ but $P(A\cap B)=1/14$, so no, they
  aren't independent. However, they are conditionally independent, in
  fact the calculations for conditional independence reduce to the
  previous calculation.


  
\item A sample space consists of the words
\begin{center}
  \{the,previous,example,had,a,mistake,in,it\}.
  \end{center}
The event $A$ consists of all words with four or fewer letters. The event $B$ consists of all words ending in a vowel. What is $P(B)$, $P(B|A)$ and $P(A|B)$?\\ \\ \\
\textbf{Solution}: So $A=\{the,had,a,in,it\}$ and $B=\{the,a,mistake,example\}$ so
$$P(B)=1/2$$ 
and
$$P(B|A)=2/5$$
and
$$P(A|B)=1/2$$

\item In the xkcd cartoon above, what is the chance the Bayesian will
  win his or her bet if the chance the sun has exploded is one in a
  million? In reality the chance is, of course, much less than one in
  a million! Show the answer to six decimal places.\\ \\ \\ \textbf{Solution}: Let $N$ be the event that the sun has exploded
  and $L$ be the event the machine says that the sun has
  exploded. Hence $P(L|N)=35/36$ whereas $P(L|\bar{N})=1/36$ and $P(N)=10^{-6}$. The Bayesian will
  win his or her bet is $P(\bar{N}|L)$:
\begin{equation}
P(\bar{N}|L)=\frac{P(L|\bar{N})P(\bar{N})}{P(L)}=\frac{1/36\times (1-10^{-6})}{1/36\times (1-10^{-6})+35/36\times 10^{-6}}=0.999965
\end{equation}


  
\end{enumerate}
  
\end{document}

