%ps1.tex
%notes for the course Probability and Statistics COMS10011 
%taught at the University of Bristol
%2019_20 Conor Houghton conor.houghton@bristol.ac.uk

%To the extent possible under law, the author has dedicated all copyright 
%and related and neighboring rights to these notes to the public domain 
%worldwide. These notes are distributed without any warranty. 


\documentclass[11pt,a4paper]{scrartcl}
\typearea{12}
\usepackage{graphicx}
%\usepackage{pstricks}
\usepackage{listings}
\usepackage{color}
\usepackage{tikz}
\usetikzlibrary{decorations.markings}
\lstset{language=C}
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{\texttt{cs-uob.github.io/COMS10014/ and github.com/coms10011/2020\_21}}
\lfoot{COMS10014 - P\&C ws1 - Conor}
\begin{document}

\section*{Probability and Combinatorics Worksheet 2}

\subsection*{Useful facts}

\begin{itemize}
\item \textbf{Combinations}: The number of ways of splitting $n$ items into sets of size $r_1$, $r_2$ through to $r_k$ with
  \begin{equation}
    r_1+r_2+\ldots+r_k=n
  \end{equation}
  is
\begin{equation}
\left(\begin{array}{c}n\\r_1,r_2,\ldots,r_k\end{array}\right)=\frac{n!}{r_1!r_2!\ldots r_k!} 
\end{equation}

\item The \textbf{conditional probability} of event $R$ given $C$:
  \begin{equation}
P(R|C)=\frac{P(R\cap C)}{P(C)} 
  \end{equation}
  This is the probability of getting an outcome in event $R$ if we
  know the outcome is in event $C$.

  \end{itemize}

\subsection*{Questions}

These are the questions you should make sure you work on in the workshop.

\begin{enumerate}

\item How many distinct anagrams has the word `BOOKKEEPER'?

\item Two events $A$ and $B$ have probabilities $P(A)=0.2$, $P(B)=0.3$ and $P(A\cup B)=0.4$. Find
\begin{enumerate}
\item Find $P(A\cap B)$.
\item Find $P(\bar{A}\cap \bar{B})$.
\item Find $P(A|B)$.
\end{enumerate}

\item In tic-tac-toe players take turns to write a X or O in a empty location in a $3\times 3$ grid, the grid is traditionally drawn like this
  {\Huge
\begin{center}
\begin{tabular}{c|c|c}
 \phantom{X}&\phantom{O}  &\phantom{O}  \\ \hline
 &\phantom{X}  & \phantom{X}\\ \hline
 &  &\phantom{O}
\end{tabular}
\end{center}
}
  X goes first so after one move the board might look like
  {\Huge
\begin{center}
  \begin{tabular}{c|c|c}
     \phantom{X}&\phantom{O}  &\phantom{O}  \\ \hline
  X&  & \\ \hline
 &  &\phantom{O}
\end{tabular}
\end{center}
}
and if the O player is very poor, or suffering from despair, the game might end after five moves with
  {\Huge
\begin{center}
  \begin{tabular}{c|c|c}
     X&O  &\phantom{O}  \\ \hline
  X&  & \\ \hline
 X&  &O
\end{tabular}
\end{center}
  } A player wins then they have a complete row, column or diagonal,
  in this case X has a column.  How many five move games of
  tic-tac-toe are there?

\item You roll a dice twice, what is the probability the second roll
  has a lower value than the first? You take the ace, two, three,
  four, five and six of hearts to make a mini-pack of cards. You draw
  two cards, what is the probability the second will be lower than the
  first, with the ace counting as a one?

  
\end{enumerate}

\subsection*{Extra questions}

These are questions you can work on in the workshop.

\begin{enumerate}

\item In a lattice path problem you have a two-dimensional lattice
  whose points are located at points with coordinates $(n,m)$ where
  $n$ and $m$ are integers. A lattice path can only have steps that
  increase either the $x$- or $y$-coordinate. This is a $5\times5$ with an example path from (0,0) to (5,5) in red.
\begin{center}
\begin{tikzpicture}
\draw (0,0) grid (5,5);
\foreach \i in {0,...,5}
    \foreach \j in {0,...,5}
    \fill (\i,\j) circle (1pt);
    \draw[ultra thick,red] (0,0) -- (0,1) -- (1,1) -- (2,1) -- (2,2)--(2,3)--(2,4)--(3,4)--(4,4)--(5,4)--(5,5);
\end{tikzpicture}
\end{center}
How many paths are there from (0,0) to (4,5)? If you pick a random path from (0,0) to (4,5) what is the probability it goes through $(2,2)$?

\item How many six move games of tic-tac-toe are there? Remember that
  a six move game can't have been won after five moves.

\item If $p$ is a probability mass function and we define a function on events:
  \begin{equation}
    P(E)=\sum_{x\in E}p(x)
  \end{equation}
  show $P$ is a probability.
  
\end{enumerate}

\subsection*{Tic-tac-tic-tac-toe}

Tic-tac-toe is a terrible game but tic-tac-tic-tac-toe is
excellent. In tic-tac-tic-tac-toe nine tic-tac-toe boards are arranged
in $3\times 3$ grid so it looks like a big tic-tac-toe board with a
little tic-tac-toe board in each of its nine squares. X can place
their first move in any of the 81 small squares, but from then the
move a player makes determines which small board the other player can
play in; so if X plays the top left square of the small board in the
middle then O must play in the small board in the top left, if they
play the middle square of that small board, then X must play in the
middle small board. When a small board is full, a player directed
there can play in any available square. The winner is the person who
wins the most small games. There is a better description on wikipedia
which calls tic-tac-tic-tac-toe `ultimate tic-tac-toe'.

\end{document}

